---
title: "Analyses"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(message=FALSE,
                      warning=FALSE)
```

# Dependencies

```{r}

library(tidyverse)
library(scales)
library(readxl)
library(janitor)
library(knitr)
library(kableExtra)

```

# Load data

```{r}

data_requests <- 
  read_excel("../data/results of data sharing requests - blinded and shuffled.xlsx", sheet = "contacted") |>
  janitor::clean_names() |>
  mutate(data_sharing_statement = ifelse(data_sharing_statement == "NA", NA, data_sharing_statement),
         data_sharing_statement_adhered_to = ifelse(data_sharing_statement_adhered_to == "NA", NA, data_sharing_statement_adhered_to),
         data_sharing_statement_type = ifelse(data_sharing_statement_type == "NA", "No data availability statement", data_sharing_statement_type))

```

# Descriptions

## N articles

N articles whose authors I contacted. I.e., all returned by existing systematic search of IRAP articles published 2018-2022 not authored by me or Chad Drake, as I already have that data.

```{r}

n <- data_requests |>
  count(name = "n articles")

n |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Articles by journal

Counts and percentages by journals

```{r}

data_requests|>
  count(journal) |>
  arrange(desc(n)) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Data sharing policies by journal

Data sharing policies for the journals who publish the most IRAP articles:

- The Psychological Record has had a [data sharing policy](https://www.springer.com/journal/40732/submission-guidelines#Instructions%20for%20Authors_Research%20Data%20Policy%20and%20Data%20Availability%20Statements) that requires data sharing except in circumstances that must be justified at time of submission: "A submission to the journal implies that materials described in the manuscript, including all relevant raw data, will be freely available to any researcher wishing to use them for non-commercial purposes, without breaching participant confidentiality... All original research must include a data availability statement."
- International Journal of Psychology & Psychological Therapy	does not have a [data sharing policy](https://www.ijpsy.com/normas.html). 
- Journal of Contextual Behavioral Science has had a [data sharing policy](https://www.elsevier.com/journals/journal-of-contextual-behavioral-science/2212-1447/guide-for-authors) that requires data sharing except in circumstances that must be justified at time of submission: "It is expected that all authors who publish in the Journal of Contextual Behavioral Science will share data upon reasonable request. Therefore, we ask authors who do not already have their data openly available to the public to include an author note indicating "Data is available upon reasonable request.". Authors can request to leave this note out if they can provide an adequately strong justification for not doing so in the cover letter."
- [data sharing policy](https://www.frontiersin.org/guidelines/policies-and-publication-ethics) that requires data sharing except in circumstances that must be justified at time of submission: "Frontiers requires that authors make the "minimal data set" underlying the findings described and used to reach the conclusions of the manuscript, available to any qualified researchers."

## Responses

Counts and percentages of articles whose authors replied to my emails, reported being able and willing to share, and actually sharing.

```{r}

data_requests |>
  pivot_longer(cols = c(replied_to_emails, reported_able_and_willing_to_share_data, shared_data),
               names_to = "metric",
               values_to = "response") |>
  group_by(metric) |>
  summarize(n = sum(response),
            percent = round_half_up(mean(response)*100, 1)) |>
  kable() |>
  kable_classic(full_width = FALSE)

```


## Prevalence of data availability statements

Count and percentages of articles with data sharing statements (overall and by year)

```{r}

data_requests |>
  mutate(data_sharing_statement = ifelse(data_sharing_statement == FALSE, FALSE, TRUE)) |>
  summarize(n_articles = n(),
            percent_data_sharing_statement = round_half_up(mean(data_sharing_statement, na.rm = TRUE)*100, 1)) |>
  kable() |>
  kable_classic(full_width = FALSE)

data_percent_data_sharing_statements <- data_requests |>
  mutate(data_sharing_statement = ifelse(data_sharing_statement == FALSE, FALSE, TRUE)) |>
  group_by(year) |>
  summarize(n_articles = n(),
            percent_data_sharing_statement = round_half_up(mean(data_sharing_statement, na.rm = TRUE)*100, 1)) 

data_percent_data_sharing_statements |>
  kable() |>
  kable_classic(full_width = FALSE)

ggplot(data_percent_data_sharing_statements, aes(year, percent_data_sharing_statement/100)) +
  geom_col() + 
  geom_text(aes(label = paste0(percent_data_sharing_statement, "%"), vjust = -0.3)) +
  ylab("Articles with data availability statement") +
  xlab("Year") +
  theme_linedraw() +
  scale_y_continuous(labels = scales::percent) +
  theme(panel.grid.minor = element_blank())
  
```

## Prevalence of data sharing

```{r}

data_percent_shared_data <- data_requests |>
  group_by(year) |>
  summarize(n_articles = n(),
            percent_shared_data = round_half_up(mean(shared_data, na.rm = TRUE)*100, 1)) 

data_percent_shared_data |>
  kable() |>
  kable_classic(full_width = FALSE)

ggplot(data_percent_shared_data, aes(year, percent_shared_data/100)) +
  geom_col() + 
  geom_text(aes(label = paste0(percent_shared_data, "%"), vjust = -0.3)) +
  ylab("Articles that shared data upon request") +
  xlab("Year") +
  theme_linedraw() +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) +
  theme(panel.grid.minor = element_blank())
  
```

## Data sharing by availability statements 

Count and percentage of articles with data sharing statement that shared data by statement type

```{r}

counts <- data_requests |>
  select(shared_data, data_sharing_statement_type) |>
  group_by(data_sharing_statement_type) |>
  summarize(n = n(),
            n_shared_data = sum(shared_data, na.rm = TRUE),
            percent_shared_data = round_half_up(mean(shared_data, na.rm = TRUE)*100, 1)) 

counts |>
  kable() |>
  kable_classic(full_width = FALSE)

```

Data was shared by some authors, but oddly no authors with promissory data sharing statements shared data when asked. I.e., data only came from actual data sharing statements and authors with no statements who nonetheless shared on request.

## Are promissory data availability statements associated with greater sharing upon request?

Ie excluding cases with data sharing statements stating actual sharing upon publication. 

Chi square test

```{r}

tab_results <- counts |>
  filter(data_sharing_statement_type != "actual") |>
  select(n, n_shared_data) |>
  as.matrix()

chisq.test(tab_results, correct = FALSE) |>
  broom::tidy() |>
  kable() |>
  kable_classic(full_width = FALSE)

```

# Session info

```{r}

sessionInfo()

```
